%!TEX TS-program = lualatex
%!TEX encoding = UTF-8 Unicode

\documentclass[11pt]{article}
\usepackage{graphicx}
	\graphicspath{{/Users/goby/Pictures/teach/438/lab/}} % set of paths to search for images

\usepackage{geometry}
\geometry{letterpaper}                   
\geometry{bottom=1in}
%\geometry{landscape}                % Activate for for rotated page geometry
\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{amssymb}
%\usepackage{mathtools}
%	\everymath{\displaystyle}

%\usepackage{color}
%\pagenumbering{gobble}

\usepackage{fontspec}
\setmainfont[Ligatures={Common, TeX}, BoldFont={* Bold}, ItalicFont={* Italic}, BoldItalicFont={* BoldItalic}, Numbers={Proportional}]{Linux Libertine O}
\setsansfont[Scale=MatchLowercase,Ligatures=TeX]{Linux Biolinum O}
\setmonofont[Scale=MatchLowercase]{Inconsolata}
\usepackage{microtype}

\usepackage{unicode-math}
\setmathfont[Scale=MatchLowercase]{Asana-Math.otf}
%\setmathfont{XITS Math}

% To define fonts for particular uses within a document. For example, 
% This sets the Libertine font to use tabular number format for tables.
%\newfontfamily{\tablenumbers}[Numbers={Monospaced}]{Linux Libertine O}
%\newfontfamily{\libertinedisplay}{Linux Libertine Display O}


\usepackage{booktabs}
\usepackage{longtable}
\usepackage{multicol}
%\usepackage{tabularx}
%\usepackage{siunitx}
%\usepackage[justification=raggedright, singlelinecheck=off]{caption}
%\captionsetup{labelsep=period} % Removes colon following figure / table number.
%\captionsetup{tablewithin=none}  % Sequential numbering of tables and figures instead of
%\captionsetup{figurewithin=none} % resetting numbers within each chapter (Intro, M&M, etc.)
%\captionsetup[table]{skip=0pt}

\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}

\usepackage{enumitem}
%\usepackage{hyperref}
%\usepackage{placeins} %P4ovides \FloatBarrier to flush all floats before a certain point.

\usepackage{titling}
\setlength{\droptitle}{-50pt}
\posttitle{\par\end{center}}
\predate{}\postdate{}

\usepackage{hanging}

\usepackage{fancyhdr}
\fancyhf{}
\pagestyle{fancy}
%\lhead{}
%\chead{}
%\rhead{Name: \rule{5cm}{0.4pt}}
%\renewcommand{\headrulewidth}{0pt}
\setlength{\headheight}{14pt}
\fancyhead[R]{\footnotesize Biogeographic Regions \thepage}
\fancyhead[L]{\footnotesize Biogeography}

\newcommand{\bigSpace}{\vspace{5\baselineskip}}

\newlength{\myLength}
\setlength{\myLength}{\parindent}

\title{Biogeographic Regions}
\author{Primary Freshwater Fishes}
\date{}                                           % Activate to display a given date or no date

\begin{document}
\maketitle
\thispagestyle{empty}


You recently learned that the world's continental landmasses are divided
into seven major biogeographic regions. \textbf{List below the seven
regions, without looking at any source}.

\begin{multicols}{2}
a.\vspace{0.5\baselineskip}

b.\vspace{0.5\baselineskip}

c.\vspace{0.5\baselineskip}

d.\vspace{0.5\baselineskip}

\columnbreak

e.\vspace{0.5\baselineskip}

f.\vspace{0.5\baselineskip}

g.\vspace{0.5\baselineskip}

\end{multicols}

The seven major regions were originally determined in large part because
mammals and birds showed similar patterns of endemism associated with
the different landmasses. Today, you will learn whether three major
groups of freshwater organisms (fishes, mussels, crayfishes) show
patterns of endemism similar to their terrestrial counterparts.

Work together in pairs to answer the following questions.

\begin{enumerate}[leftmargin=*]
\item Do you think the three freshwater groups will show patterns of
endemism associated with the seven biogeographic regions? Explain why or
why not.\vspace{10\baselineskip}

\end{enumerate}

You were lucky enough to draw primary freshwater fishes for this
exercise. Primary freshwater fishes cannot tolerate salt water. You have
been given a table of data that lists the orders and families of primary
freshwater fishes from around the world, and indicates whether the taxon
is present (1) or absent (0) in each of the biogeographic regions.

\begin{enumerate}[resume, leftmargin=*]
\item How many orders of fishes are found in each biogeographic region?

\begin{multicols}{2}
a.\vspace{0.5\baselineskip}

b.\vspace{0.5\baselineskip}

c.\vspace{0.5\baselineskip}

d.\vspace{0.5\baselineskip}

e.\vspace{0.5\baselineskip}

f.\vspace{0.5\baselineskip}

g.\vspace{0.5\baselineskip}
\end{multicols}

\item  How many families of fishes are found in each biogeographic
region?


\begin{multicols}{2}
a.\vspace{0.5\baselineskip}

b.\vspace{0.5\baselineskip}

c.\vspace{0.5\baselineskip}

d.\vspace{0.5\baselineskip}

e.\vspace{0.5\baselineskip}

f.\vspace{0.5\baselineskip}

g.\vspace{0.5\baselineskip}
\end{multicols}

\item Which region has the greatest number of orders? Which region has the
least number of orders? Which region has the greatest number of
families? The least? Are they the same regions you listed for the
orders?

\begin{multicols}{2}

Greatest number of orders: \vspace{2\baselineskip}

Least number of orders: \vspace{2\baselineskip}

 Greatest number of families: \vspace{2\baselineskip}
 
 Least number of families: \vspace{2\baselineskip}
 
 \end{multicols}
 
 \vspace{\baselineskip}
 
\item List the families that are found in more than one biogeographic region.
Name the regions for each family you list.\vspace*{\stretch{1}}

\item Do any families show a Pangaea, Laurasian, or Gondwanan distribution?
If so, list the families, and then explain for each family how you
determined which distribution pattern it has.\vspace*{\stretch{1}}

\newpage
\item Do any of the orders show a Pangaean, Laurasian or Gondwanan
distribution? If so, list the orders, and then explain for each order
how you determined which distribution pattern it has.\vspace*{\stretch{1}}

\item Are there any orders or families with a distribution that is
\emph{not apparently} consistent with plate tectonics, i.e., a
distribution that is not Pangaean, Laurasian or Gondwanan? If so, list
the taxa and regions in which the taxa are located. How might you try to
explain this distribution? (Time to think about the biology of the
organisms.)\vspace*{\stretch{1}}

\end{enumerate}

\newpage

The data sheet includes a distribution map and two phylogenetic
hypotheses for the family Cichlidae. Study the distribution and the
phylogenetic trees carefully, then answer the following question.

\begin{enumerate}[resume, leftmargin=*]
\item Can you tell from the phylogeny the order of the breakup of the
continents that composed Gondwana? That is, based on the phylogeny of
the cichlids, which continental landmasses do you think separated first
from Gondwana? Which separated last? Which breakups were somewhere
between the others? Explain.\vspace*{\stretch{1}}

\end{enumerate}

Be prepared to come to the front of the class to discuss some of your
results.


\end{document}  