%!TEX TS-program = lualatex
%!TEX encoding = UTF-8 Unicode

\documentclass[11pt]{article}
\usepackage{graphicx}
	\graphicspath{{/Users/goby/Pictures/teach/438/lab/}} % set of paths to search for images

\usepackage{geometry}
\geometry{letterpaper}                   
\geometry{bottom=1in}
%\geometry{landscape}                % Activate for for rotated page geometry
\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{amssymb}
%\usepackage{mathtools}
%	\everymath{\displaystyle}

%\usepackage{color}
%\pagenumbering{gobble}

\usepackage{fontspec}
\setmainfont[Ligatures={Common, TeX}, BoldFont={* Bold}, ItalicFont={* Italic}, BoldItalicFont={* BoldItalic}, Numbers={Proportional}]{Linux Libertine O}
\setsansfont[Scale=MatchLowercase,Ligatures=TeX]{Linux Biolinum O}
\setmonofont[Scale=MatchLowercase]{Inconsolata}
\usepackage{microtype}

\usepackage{unicode-math}
\setmathfont[Scale=MatchLowercase]{Asana-Math.otf}
%\setmathfont{XITS Math}

% To define fonts for particular uses within a document. For example, 
% This sets the Libertine font to use tabular number format for tables.
%\newfontfamily{\tablenumbers}[Numbers={Monospaced}]{Linux Libertine O}
%\newfontfamily{\libertinedisplay}{Linux Libertine Display O}


\usepackage{booktabs}
\usepackage{longtable}
\usepackage{multicol}
%\usepackage{tabularx}
%\usepackage{siunitx}
%\usepackage[justification=raggedright, singlelinecheck=off]{caption}
%\captionsetup{labelsep=period} % Removes colon following figure / table number.
%\captionsetup{tablewithin=none}  % Sequential numbering of tables and figures instead of
%\captionsetup{figurewithin=none} % resetting numbers within each chapter (Intro, M&M, etc.)
%\captionsetup[table]{skip=0pt}

\usepackage{array}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}

\usepackage{enumitem}
%\usepackage{hyperref}
%\usepackage{placeins} %P4ovides \FloatBarrier to flush all floats before a certain point.

\usepackage{titling}
\setlength{\droptitle}{-50pt}
\posttitle{\par\end{center}}
\predate{}\postdate{}

\usepackage{hanging}

\usepackage{fancyhdr}
\fancyhf{}
\pagestyle{fancy}
%\lhead{}
%\chead{}
%\rhead{Name: \rule{5cm}{0.4pt}}
%\renewcommand{\headrulewidth}{0pt}
\setlength{\headheight}{14pt}
\fancyhead[R]{\footnotesize Biogeographic Regions \thepage}
\fancyhead[L]{\footnotesize Biogeography}

\newcommand{\bigSpace}{\vspace{5\baselineskip}}

\newlength{\myLength}
\setlength{\myLength}{\parindent}

\title{Biogeographic Regions}
\author{Crayfishes}
\date{}                                           % Activate to display a given date or no date

\begin{document}
\maketitle
\thispagestyle{empty}


You recently learned that the world's continental landmasses are divided
into seven major biogeographic regions. \textbf{List below the seven
regions, without looking at any source}.

\begin{multicols}{2}
a.\vspace{0.5\baselineskip}

b.\vspace{0.5\baselineskip}

c.\vspace{0.5\baselineskip}

d.\vspace{0.5\baselineskip}

\columnbreak

e.\vspace{0.5\baselineskip}

f.\vspace{0.5\baselineskip}

g.\vspace{0.5\baselineskip}

\end{multicols}

The seven major regions were originally determined in large part because
mammals and birds showed similar patterns of endemism associated with
the different landmasses. Today, you will learn whether three major
groups of freshwater organisms (fishes, mussels, crayfishes) show
patterns of endemism similar to their terrestrial counterparts.

Work together in pairs to answer the following questions.

\begin{enumerate}[leftmargin=*]
\item Do you think the three freshwater groups will show patterns of
endemism associated with the seven biogeographic regions? Explain why or
why not.\vspace{10\baselineskip}

\end{enumerate}

You were lucky enough to draw crayfishes for this exercise. You have
been given a sheet of data that lists the superfamilies, families and
genera of crayfishes from around the world, and the number of species
for each genus found in each biogeographic region.

\begin{enumerate}[resume, leftmargin=*]
\item How many genera of crayfishes are found in each biogeographic region?

\begin{multicols}{2}
a.\vspace{0.5\baselineskip}

b.\vspace{0.5\baselineskip}

c.\vspace{0.5\baselineskip}

d.\vspace{0.5\baselineskip}

e.\vspace{0.5\baselineskip}

f.\vspace{0.5\baselineskip}

g.\vspace{0.5\baselineskip}
\end{multicols}

\item  How many species of crayfishes are found in each biogeographic
region?


\begin{multicols}{2}
a.\vspace{0.5\baselineskip}

b.\vspace{0.5\baselineskip}

c.\vspace{0.5\baselineskip}

d.\vspace{0.5\baselineskip}

e.\vspace{0.5\baselineskip}

f.\vspace{0.5\baselineskip}

g.\vspace{0.5\baselineskip}
\end{multicols}

\item Which region has the greatest number of genera? Which region has the
fewest number of genera? Which region has the greatest number of
species? The fewest? Are they the same regions you listed for the
genera?

\begin{multicols}{2}

%\vspace*{\baselineskip}

Greatest number of genera:\vspace{2\baselineskip}

Least number of genera: \vspace{2\baselineskip}

 Greatest number of species:\vspace{2\baselineskip}
 
 Least number of species:\vspace{2\baselineskip}
 
 \end{multicols}
 
 \vspace{\baselineskip}
 
\item List the genera that are found in more than one biogeographic region.
Name the regions for each genus you list.\vspace*{\stretch{1}}

\item Do any genera show a Pangaea, Laurasian, or Gondwanan distribution?
If so, list the genera, and then explain for each genus how you
determined which distribution pattern it has.\vspace*{\stretch{1}}

\newpage
\item Do any of the families show a Pangaean, Laurasian or Gondwanan
distribution? If so, list the families, and then explain for each family
how you determined which distribution pattern it has.\vspace*{\stretch{1}}

\item Are there any families or genera with a distribution that is
\emph{not apparently} consistent with plate tectonics, i.e., a
distribution that is not Pangaean, Laurasian or Gondwanan? If so, list
the taxa and regions in which the taxa are located. How might you try to
explain this distribution?\vspace*{\stretch{1}}

\item Centers of diversity are areas where species diversity is especially
high. Some biologists think that these areas are places where new
species evolve frequently within the higher taxon. Based on the
information you compiled above, where do you think the center of
diversity is for the Astacoidea? Where is the center of diversity for
the Parastacoidea?\vspace*{\stretch{1}}

\end{enumerate}

\newpage

Freshwater crayfishes evolved from marine clawed lobster ancestors.
Because each superfamily has a different center of diversity, a
hypothesis was proposed in 1880, which stated that freshwater crayfish 
evolved from marine ancestors twice, once for each superfamily. In 1902,
a competing hypothesis proposed that freshwater crayfish evolved only
once.

Carefully study the two phylogenetic trees that were included with the
distribution data, then answer the next two questions.

\begin{enumerate}[resume, leftmargin=*]
\item The phylogenetic trees clearly support one of the two hypotheses.
Which one? Explain.\vspace*{\stretch{1}}

\item Is the hypothesized relationship among the higher crayfish taxa
consistent with the breakup of Pangaea? Explain.\vspace*{\stretch{1}}

\end{enumerate}

Be prepared to come to the front of the class to discuss some of your
results.


\end{document}  